---
title: Regression analysis for each word
output:
  html_document:
    highlight: tango
    toc: true
    number_sections: false
    theme: cerulean
    toc_float: true
    code_folding: hide
---

```{r setup, message=FALSE, warning = F}
library(knitr)

opts_chunk$set(echo = T, message = F, warning = F, 
               error = F, cache = F, tidy = F)

library(tidyverse)
library(feather)
library(langcog)
library(modelr)
library(broom)
library(corrplot)


theme_set(theme_classic(base_size = 10))

```


## Predicting mtld at t2
MODEL: `lm(log_mtld_t2 ~ know_word_at_t1 + log_mtld_t1 + age_t1 + age_diff + log(n_transcripts_t1) + log(n_transcripts_t2), complete_df)`

```{r}
word_coeffs_min5_t2 <- read_csv("data/word_coeffs_log_mtld_t2.csv") %>%
  mutate(word = tolower(word))

ggplot(word_coeffs_min5_t2, aes(t)) +
  geom_histogram() +
  ggtitle("t-distribution ") +
  geom_vline(aes(xintercept = 2), color = "red") +
  geom_vline(aes(xintercept = -2), color = "red") +
  theme_classic()

```

```{r}
 word_coeffs_min5_t2 %>%
  arrange(-t) %>%
  DT::datatable()
```


## Sanity Check

Do kids who have a high mean t have high mtld at t2? Yes.
```{r}
all_types <- read_csv("/Users/mollylewis/Documents/research/Projects/1_in_progress/VOCAB_SEEDS/analyses/1_mtld_measure/data/target_types_for_MTLD_kids_600_900.csv") 
MINWORDSFORVOCAB <- 5
word_counts <- all_types %>%
  filter(tbin == "t1") %>%
  mutate(gloss_clean = tolower(gloss))   %>%
  group_by(target_child_id, gloss_clean) %>%
  summarize(count = sum(count)) %>%
  filter(count >= MINWORDSFORVOCAB)

t1_word_counts_with_ts <- word_counts  %>%
  left_join(word_coeffs_min5_t2 %>% select(word, t), 
            by = c("gloss_clean" = "word")) %>%
  select(-gloss_clean, -count) %>%
  group_by(target_child_id) %>%
  summarize(sum_t = sum(t, na.rm = T),
            mean_t = mean(t, na.rm = T)) 

mtld_age <- read_csv("/Users/mollylewis/Documents/research/Projects/1_in_progress/VOCAB_SEEDS/analyses/1_mtld_measure/data/groups_info_600_900_corrected.csv") %>%
    mutate(log_mtld_t1 = log(mtld_t1),
         log_mtld_t2 = log(mtld_t2),
         age_diff = age_t2 - age_t1) %>%
    select(target_child_id, log_mtld_t1, log_mtld_t2, 
           age_t1, age_t2, age_diff, corpus_name)

t1_word_counts_with_ts_mtld <- t1_word_counts_with_ts %>%
  left_join(mtld_age) %>%
  select(-target_child_id)

t1_word_counts_with_ts_mtld %>%
ggplot( aes(x =  mean_t , y = log_mtld_t2)) +
  geom_point()  +
  geom_smooth(method = "lm") +
  theme_classic()
```

Does it hold controlling for stuff? Yes.
```{r}
transcript_length <- read_csv("/Users/mollylewis/Documents/research/Projects/1_in_progress/VOCAB_SEEDS/analyses/3_kid_vocabs/data/transcript_length_by_kid.csv") %>%
  mutate(log_transcript_length_t1 = log(transcript_length_t1),
         log_transcript_length_t2 = log(transcript_length_t2)) %>%
  select(-transcript_length_t1, transcript_length_t2)

freq_info <- read_csv("/Users/mollylewis/Documents/research/Projects/1_in_progress/VOCAB_SEEDS/analyses/3_kid_vocabs/data/frequency_based_on_input_by_kid.csv")

t1_word_counts_with_ts_mtld <- t1_word_counts_with_ts %>%
  left_join(mtld_age) %>%
  left_join(transcript_length) %>%
  left_join(freq_info)

lm(log_mtld_t2 ~ log_mtld_t1 + mean_t + log_transcript_length_t1 + log_transcript_length_t2 + mean_freq_t1 + mean_freq_t2 + age_diff + age_t1  , t1_word_counts_with_ts_mtld  ) %>%
  summary()
```

## Predicting t-value of a word with other measures
```{r}
freq <- read_csv("/Users/mollylewis/Documents/research/Projects/1_in_progress/VOCAB_SEEDS/analyses/3_kid_vocabs/data/childes_adult_word_freq.csv")

density_norms <- read_csv("/Users/mollylewis/Documents/research/Projects/1_in_progress/VOCAB_SEEDS/analyses/3_kid_vocabs/data/bills_density_norms.csv")

embedding_dist <- read_csv("data/childes_embedding_dist_by_word.csv")

concreteness <- read_csv("/Users/mollylewis/Documents/research/Projects/2_published/ref_complex/corpus/brysbaert_database/brysbaert_corpus.csv") %>%
  rename(word = Word) %>%
  select(word, Conc.M)

concepts <- read_tsv("/Users/mollylewis/Documents/research/Projects/1_in_progress/VOCAB_SEEDS/analyses/3_kid_vocabs/data/CONCS_brm.txt") %>%
  select(Concept, Familiarity, Length_Syllables, Bigram, 14:33) %>%
  mutate(Concept = tolower(Concept),
         Concept = map_chr(Concept, ~ pluck(str_split(., "_"),1,1))) %>%
  rename(word = Concept) %>%
  select(word, Mean_Distinct_No_Tax)

word_coeffs_min5_t2_with_vars <- word_coeffs_min5_t2 %>%
  mutate(word = tolower(word)) %>%
  left_join(density_norms) %>%
  left_join(freq) %>%
  left_join(embedding_dist) %>%
  left_join(concepts) %>%
  left_join(concreteness)

lm(t ~ centrality + log_freq, word_coeffs_min5_t2_with_vars) %>%
  summary()

lm(t ~ density  + log_freq, word_coeffs_min5_t2_with_vars) %>%
  summary()

lm(t ~ mean_dist + log_freq, word_coeffs_min5_t2_with_vars) %>%
  summary()

lm(t ~ mean_dist  * log_freq, word_coeffs_min5_t2_with_vars) %>%
  summary()

lm(t ~ Conc.M +  log_freq, word_coeffs_min5_t2_with_vars) %>%
  summary()

lm(t ~ mean_dist +  Conc.M + log_freq, word_coeffs_min5_t2_with_vars) %>%
  summary()

lm(t ~ Conc.M  + mean_dist *  log_freq, word_coeffs_min5_t2_with_vars) %>%
  summary()
```

Concreteness and mean distance both predict t score - words that are similiar to other words and highly concrete have large t-scores.
