---
title: Semantic space by-kid analyses
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: tango
    toc: true
    number_sections: false
    theme: cerulean
    toc_float: false
    code_folding: hide
---

```{r setup, message = F, warning = F}
library(knitr)

opts_chunk$set(echo = T, message = F, warning = F, 
               error = F, cache = F, tidy = F)

library(tidyverse)
library(langcog)
library(data.table)
library(feather)

theme_set(theme_classic(base_size = 10))
```


```{r, params}
MINWORDSFORVOCAB <- 5
```

The min words for vocab here is `r MINWORDSFORVOCAB`.

```{r}
all_types <- read_csv("../1_mtld_measure/data/target_types_for_MTLD_kids_600_900.csv") 
mlu_info <- read_csv("mlu_by_kid.csv")
trans_info <- read_csv("t1_transitional_probs_in_vocab_missing0.csv")
mcrae_info <- read_csv("mcrae_vocab_by_kid_t1.txt") %>%
  rename(Per_Corred_Pairs_No_Tax = `%_Corred_Pairs_No_Tax`)
pos_info <- read_csv("prop_pos_by_kid_t1.csv")
conc_info <- read_csv("conc_by_kid_t1.csv")

kid_info <- read_csv("/Users/mollylewis/Documents/research/Projects/1_in_progress/VOCAB_SEEDS/analyses/4_semantic_density/semantic_density_df.csv") %>%
  left_join(mlu_info) %>%
  left_join(trans_info) %>%
  left_join(mcrae_info)%>%
  left_join(pos_info) %>%
  left_join(conc_info) %>%
  mutate(corpus_id = as.factor(corpus_id))

density <- read_csv("bills_density_norms.csv")

trigrams <- read_csv("/Users/mollylewis/Documents/research/Projects/1_in_progress/VOCAB_SEEDS/analyses/2_trigrams/data/trigrams/kid_childes_trigrams_turns_BY_KID.csv") %>%
  mutate(log_n_adult_trigrams = log(n_adult_trigrams + 1)) %>%
  select(target_child_id, log_n_adult_trigrams, tbin) %>%
  spread("tbin", "log_n_adult_trigrams") %>%
  mutate(log_n_adult_trigrams_t1 = t1,
         log_n_adult_trigrams_t2 = t2)

#freq <- read_tsv("/Users/mollylewis/Documents/research/Projects/1_in_progress/VOCAB_SEEDS/analyses/1_mtld_measure/data/control_variables/SUBTLEXus_corpus.txt") %>%
 # rename(word = Word,
  #       log_freq = Lg10WF) %>%
 # select(word, log_freq)

freq <- read_csv("childes_adult_word_freq.csv")
```

Embedding models
```{r}
word2vec_model_childes <- read_csv("3_train_childes_model/childes_adult_w2v.txt") %>%
  rename(target_word = word)
word2vec_model_wiki <- read_feather("fast_text_childes_words_600_900.feather") 
```

Get filtered version of types for each kid
```{r}
types_clean <- all_types %>%
  filter(tbin == "t1") %>%
  mutate(gloss_clean = tolower(gloss))   %>%
  group_by(target_child_id, gloss_clean) %>%
  summarize(count = sum(count)) %>%
  filter(count >= MINWORDSFORVOCAB) %>%
  mutate(log_count = log(count)) %>%
  select(-count) %>%
  left_join(freq, by= c("gloss_clean" = "word")) %>%
  mutate(log_count_w1 = log_count,
         log_count_w2 = log_count,
         log_freq_w1 = log_freq, 
         log_freq_w2 = log_freq) %>%
  select(-log_count, -log_freq)
```

```{r}
conc <- read_csv("/Users/mollylewis/Documents/research/Projects/2_published/ref_complex/corpus/brysbaert_database/brysbaert_corpus.csv") %>%
  select(Word, Conc.M)

distinct_types <- types_clean %>%
  left_join(conc, by =c("gloss_clean" = "Word")) %>%
  ungroup() %>%
  distinct(gloss_clean, .keep_all = T)

median_conc <- median(distinct_types$Conc.M, na.rm = T)

types_clean <- types_clean %>%
    left_join(conc, by =c("gloss_clean" = "Word")) %>%
    filter(Conc.M <= median_conc)
```

 
Get distance measure by kid at t1 in both wikipedia model and adult childes model
```{r}


get_vocab_measure_by_kid3 <- function(id, data, model){
  this_kids_model <- model %>%
    filter(target_word %in% data$gloss_clean)
  
  words_in_model <- data %>%
    filter(gloss_clean %in% this_kids_model$target_word)

   # get pairwise distances
   word_word_dists <- coop::cosine(t(this_kids_model[,-1])) 
   
   resid_df <- data.frame(word = this_kids_model$target_word,
                          mean_dist = rowMeans(word_word_dists)) %>%
                    left_join(freq) %>%
                    left_join(density)
   
     mean_resid <- NA
     try({
      model <- lm(mean_dist ~ log_freq, resid_df)
      resid_df_with_resids <- resid_df %>%
                modelr::add_residuals(model, "dist_resids")
      mean_resid <- mean(resid_df_with_resids$dist_resids, na.rm = T)
     })
     
     
     mean_centrality_resid <- NA
     try({
      model <- lm(centrality ~ log_freq, resid_df)
      resid_df_with_resids <- resid_df %>%
                modelr::add_residuals(model, "dist_resids")
      mean_centrality_resid <- mean(resid_df_with_resids$dist_resids, na.rm = T)
     })
                          

     data.frame(target_child_id = id,
                mean_dist_t1 = mean(log(word_word_dists + 1)),
                mean_centrality = mean_centrality_resid,
                mean_dist_resid = mean_resid, 
              #  var_dist_t1 =  ifelse(mean(var(log(word_word_dists + 1)) == 0, NA,
              #                         mean(var(log(word_word_dists + 1))))),
                n_t1 = nrow(word_word_dists),
                mean_freq_t1 =  mean(words_in_model$log_freq_w1, na.rm = T))
     
    
}

nested_data_by_kid <- nest(types_clean, -target_child_id)

vocab_measures_wiki <- map2_df(nested_data_by_kid$target_child_id, 
                          nested_data_by_kid$data, 
                          get_vocab_measure_by_kid3, 
                          word2vec_model_wiki) %>%
  rename(mean_dist_t1_wiki = mean_dist_t1, 
         n_t1_wiki = n_t1,
         mean_freq_t1_wiki = mean_freq_t1,
         mean_centrality_wiki = mean_centrality,
         mean_density_wiki = mean_density,

        mean_dist_resid_wiki= mean_dist_resid)


vocab_measures_childes <- map2_df(nested_data_by_kid$target_child_id, 
                          nested_data_by_kid$data, 
                          get_vocab_measure_by_kid3, 
                          word2vec_model_childes) %>%
  rename(mean_dist_t1_childes = mean_dist_t1, 
         n_t1_childes = n_t1,
              mean_centrality_childes = mean_centrality,
         mean_density_childes = mean_density,
         mean_freq_t1_childes  = mean_freq_t1,
         mean_dist_resid_childes = mean_dist_resid)

vocab_measures <- left_join(vocab_measures_wiki, vocab_measures_childes) 
```

### Mean distance in childes vs. mean distance in wiki model at t1
```{r}
ggplot(vocab_measures, aes(x = mean_dist_t1_childes, y = mean_dist_t1_wiki)) +
  geom_point() +
  geom_smooth(method = "lm")
```

```{r}
embedding_corr  <- cor.test(vocab_measures$mean_dist_t1_childes, vocab_measures$mean_dist_t1_wiki)
embedding_corr
```

They're highly correlated r = `r as.numeric(embedding_corr$estimate)`. 


## Other kid-level variables at t1

Get merged dataframe
```{r}
vocab_df <- vocab_measures %>%
  #  filter(n_t1_wiki > 2) %>% # can't look at median distance when there's only two words
    left_join(kid_info %>% select(target_child_id, 
                                  log_mtld_t1, 
                                  log_mtld_t2, 
                                  age_t1, age_diff , 
                                  log_transcript_length_t1, log_transcript_length_t2, mlu_m_t1, mlu_m_t2, 
                                  target_child_sex,
                                  conc_t1,
                                #  n_clusters, 
                                 # mean_trans_prob_t1 , 
                                 # prop_na_trans_t1,
                                  Per_Corred_Pairs_No_Tax,
                                  Num_Corred_Pairs_No_Tax, 
                                  prop_noun_t1, 
                                  prop_verb_t1,
                                  corpus_id, collection_name,
                                age_t2))  %>%
  left_join(trigrams %>% select(target_child_id, log_n_adult_trigrams_t1, log_n_adult_trigrams_t2))


```

descriptive stats
```{r, include = F}

DAYSINMONTH <-  30.42
vocab_df %>%
  filter(!is.na(Num_Corred_Pairs_No_Tax),
         !is.na(mlu_m_t1))  %>%
  summarize(mtld_t1 = mean(exp(log_mtld_t1)),
            mtld_t2 = mean(exp(log_mtld_t2)),
            mlu_t1 = mean(exp(mlu_m_t1)),
            mlu_t2 = mean(exp(mlu_m_t2)),
          #  log_num_trigrams_t1 = mean(exp(log_num_trigrams_t1)),
          #  log_num_trigrams_t2 = mean(exp(log_num_trigrams_t2)),
            
            min_age_t1 = min(age_t1)/DAYSINMONTH,
            min_age_t1 = min(age_t1)/DAYSINMONTH,
            max_age_t1 = max(age_t1)/DAYSINMONTH,
            age_t1 = mean(age_t1)/DAYSINMONTH,
            min_age_t2 = min(age_t2)/DAYSINMONTH,
            max_age_t2 = max(age_t2)/DAYSINMONTH,
            age_t2 = mean(age_t2)/DAYSINMONTH) %>%
  kable()

```

### Correlation between variables
```{r, fig.width=10, fig.height=10}
df_no_corrs <- vocab_df %>%
  select(-target_child_id, -target_child_sex, -corpus_id, -collection_name) 
  #filter_all(all_vars(!is.na(.)))

corr_mat <- cor(df_no_corrs, 
                use = "pairwise.complete.obs")

p.mat <- corrplot::cor.mtest(df_no_corrs, 
                  conf.level = .95,  
                  use = "pairwise.complete.obs")$p

cols <- rev(colorRampPalette(c("red", "white", "blue"))(100))

corrplot::corrplot(corr_mat, method = "color",  col = cols,
         type = "full", order = "original", number.cex = .7, #order = "hclust"
         addCoef.col = "black", insig = "blank",
         p.mat = p.mat, sig.level = .05, 
         tl.col = "black", tl.srt = 90,
         diag = FALSE)

#age_t1 
#log_transcript_length_t1
#mlu_m_t1*
#mean_trans_prob_t1
#prop_noun_t1
#median_freq_t1_wiki
```

## Predicting t2 measures
### Wikipedia{.tabset}

#### MTLD
```{r}
lme4::lmer(scale(log_mtld_t2) ~ 
                  #scale(mean_dist_resid_wiki) +
                  scale(Num_Corred_Pairs_No_Tax) +
                  scale(age_t1) + 
                  scale(prop_noun_t1) +
                  scale(log_mtld_t1) +
                  scale(mean_freq_t1_wiki) +
                  scale(mlu_m_t1) +
                  scale(log_transcript_length_t1) + 
             (1|corpus_id),
   data = filter(vocab_df, !is.na(Num_Corred_Pairs_No_Tax))) %>%
  summary()
```

#### MLU
```{r}

 lme4::lmer(scale(mlu_m_t2) ~ 
                  #scale(mean_dist_resid_wiki) +
                  scale(Num_Corred_Pairs_No_Tax) +
                                scale(mean_freq_t1_wiki) +

                  scale(age_t1) + 
                  scale(prop_noun_t1) +
                  scale(log_mtld_t1) +
                  scale(mlu_m_t1) +
                  scale(log_transcript_length_t1) + 
             (1|corpus_id),
   data = filter(vocab_df, !is.na(Num_Corred_Pairs_No_Tax))) %>%
  summary()
```
 
#### Trigrams
```{r}
  lme4::lmer(scale(log_n_adult_trigrams_t2) ~ 
                  #scale(mean_dist_resid_wiki) +
                  scale(Num_Corred_Pairs_No_Tax) +
                 # scale(mean_centrality_childes) +
                  scale(age_t1) + 
                  scale(prop_noun_t1) +
                  scale(log_n_adult_trigrams_t1) +
                  scale(mean_freq_t1_wiki) +
                  scale(mlu_m_t1) +
                  scale(log_mtld_t1) +
                  scale(log_transcript_length_t1) + 
             (1|corpus_id),
   data = filter(vocab_df, !is.na(Num_Corred_Pairs_No_Tax))) %>%
  summary()

  lme4::lmer(scale(log_n_adult_trigrams_t2) ~ 
                  #scale(mean_dist_resid_wiki) +
                  scale(Num_Corred_Pairs_No_Tax) +
                  #scale(mean_density_childes) +
                  scale(age_t1) + 
                  #scale(prop_noun_t1) +
                  scale(log_n_adult_trigrams_t1) +
                 scale(mean_freq_t1_wiki) +
                  scale(mlu_m_t1) +
                  scale(log_mtld_t1) +
                  scale(log_transcript_length_t1) + 
             (1|corpus_id),
   data = vocab_df) %>%
  summary()
  
  
  
```

### Childes{.tabset}


#### MTLD
```{r}
lme4::lmer(scale(log_mtld_t2) ~ 
                  #scale(mean_dist_resid_childes) +
                  scale(mean_centrality_childes) +
                  scale(Num_Corred_Pairs_No_Tax) +
                  scale(age_t1) + 
                  scale(prop_noun_t1) +
                  scale(log_mtld_t1) +
                  scale(mlu_m_t1) +
                  #scale(mean_freq_t1_childes) +
                  scale(log_transcript_length_t1) + 
             (1|corpus_id),
   data = filter(vocab_df, !is.na(Num_Corred_Pairs_No_Tax))) %>%
  summary()

lme4::lmer(scale(log_mtld_t2) ~ 
                  scale(mean_dist_resid_childes) +
                  #scale(Num_Corred_Pairs_No_Tax) +
                  scale(age_t1) + 
                  scale(prop_noun_t1) +
                  scale(log_mtld_t1) +
                  scale(mlu_m_t1) +
                  scale(log_transcript_length_t1) + 
             (1|corpus_id),
   data =vocab_df) %>%
  summary()


```

#### MLU
```{r}

 lme4::lmer(scale(mlu_m_t2) ~ 
                  scale(mean_dist_resid_childes) +
                  scale(Num_Corred_Pairs_No_Tax) +
                  scale(age_t1) + 
                  scale(prop_noun_t1) +
                  scale(log_mtld_t1) +
                  scale(mlu_m_t1) +
                  scale(log_transcript_length_t1) + 
             (1|corpus_id),
   data = filter(vocab_df, !is.na(Num_Corred_Pairs_No_Tax))) %>%
  summary()
```

#### Trigrams
```{r}
 
  lme4::lmer(scale(log_num_trigrams_t2) ~ 
                  #scale(mean_dist_resid_childes) +
                  scale(mean_centrality_childes) +

                  scale(Num_Corred_Pairs_No_Tax) +
                  scale(age_t1) + 
                  scale(prop_noun_t1) +
                  scale(log_num_trigrams_t1) +
                  scale(mlu_m_t1) +
                  scale(log_mtld_t1) +
                                 scale(mean_freq_t1_childes) +

                  scale(log_transcript_length_t1) + 
             (1|corpus_id),
   data = filter(vocab_df, !is.na(Num_Corred_Pairs_No_Tax))) %>%
  summary()
```


## Spaghetti plot
```{r}
plot_df <- vocab_df %>%
  filter(!is.na(Num_Corred_Pairs_No_Tax)) %>%
  filter(!is.na(mlu_m_t1))

OUTFILE <- "spaghetti_plot.pdf"
#pdf(OUTFILE)
plot_df %>%
  rename(`Time 1` = "log_mtld_t1",
         `Time 2` = "log_mtld_t2") %>%
  select(`Time 1`, `Time 2`, target_child_id) %>%
  gather("timepoint", "value", -target_child_id) %>%
  ggplot(aes(x = timepoint, y = value, group = target_child_id)) +
  ylab("Vocabulary size estimate (log MTLD)") +
  xlab("") +
  theme(axis.line.x = element_blank(),
        axis.line.y = element_line(size = 1),
        axis.ticks = element_blank(),
        axis.title.y = element_text(size = 14),
        axis.text.x = element_text(size = 14, color = "black"),
        axis.text.y = element_text(size = 12)) + 
  geom_line() +
  geom_point(color = "red") 
#dev.off()
```


### Comparing frequency measures
```{r, include = F, eval = F}

freq1 <- read_tsv("/Users/mollylewis/Documents/research/Projects/1_in_progress/VOCAB_SEEDS/analyses/1_mtld_measure/data/control_variables/SUBTLEXus_corpus.txt") %>%
  rename(word = Word,
         log_freq_sublex = Lg10WF) %>%
  select(word, log_freq_sublex)

freq2 <- read_csv("childes_adult_word_freq.csv")

aoa <- read_csv("/Users/mollylewis/Downloads/kuperman_aoa.csv") %>%
  select(Word, Rating.Mean) %>%
    rename(word = Word,
         aoa = Rating.Mean) 

all_freqs <- freq2 %>%
  left_join(freq1) %>%
  left_join(aoa) %>%
  filter(!is.na(log_freq_sublex))

cor.test(all_freqs$aoa, all_freqs$log_freq)
cor.test(all_freqs$aoa, all_freqs$log_freq_sublex)

ggplot(all_freqs, aes(x =log_freq, y = aoa)) +
  geom_point() +
  geom_smooth(method = "lm") 

ggplot(all_freqs, aes(x = log_freq_sublex, y = aoa)) +
  geom_point() +
  geom_smooth(method = "lm")



types_clean %>%
  filter(target_child_id %in% plot_df$target_child_id) %>%
  ggplot(aes(x = log_freq_w2)) +
  geom_histogram() +
  facet_wrap(~target_child_id)
```