---
title: Semantic space by-kid analyses
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: tango
    toc: true
    number_sections: false
    theme: cerulean
    toc_float: false
    code_folding: hide
---

```{r setup, message = F, warning = F}
library(knitr)

opts_chunk$set(echo = T, message = F, warning = F, 
               error = F, cache = F, tidy = F)

library(tidyverse)
library(langcog)
library(data.table)
library(feather)

theme_set(theme_classic(base_size = 10))
```


```{r, params}
MINWORDSFORVOCAB <- 5
```

The min words for vocab here is `r MINWORDSFORVOCAB`.

```{r}
all_types <- read_csv("../1_mtld_measure/data/target_types_for_MTLD_kids_600_900.csv") 
mlu_info <- read_csv("mlu_by_kid.csv")
trans_info <- read_csv("t1_transitional_probs_in_vocab_missing0.csv")
mcrae_info <- read_csv("mcrae_vocab_by_kid_t1.txt") %>%
  rename(Per_Corred_Pairs_No_Tax = `%_Corred_Pairs_No_Tax`)
pos_info <- read_csv("prop_pos_by_kid_t1.csv")
conc_info <- read_csv("conc_by_kid_t1.csv")

kid_info <- read_csv("/Users/mollylewis/Documents/research/Projects/1_in_progress/VOCAB_SEEDS/analyses/4_semantic_density/semantic_density_df.csv") %>%
  left_join(mlu_info) %>%
  left_join(trans_info) %>%
  left_join(mcrae_info)%>%
  left_join(pos_info) %>%
  left_join(conc_info) %>%
  mutate(corpus_id = as.factor(corpus_id))

freq <- read_tsv("/Users/mollylewis/Documents/research/Projects/1_in_progress/VOCAB_SEEDS/analyses/1_mtld_measure/data/control_variables/SUBTLEXus_corpus.txt") %>%
  rename(word = Word,
         log_freq = Lg10WF) %>%
  select(word, log_freq)
```

```{r}

```

Embedding models
```{r}
word2vec_model_childes <- read_csv("3_train_childes_model/childes_adult_w2v.txt") %>%
  rename(target_word = word)
word2vec_model_wiki <- read_feather("fast_text_childes_words_600_900.feather") 
```

Get filtered version of types for each kid
```{r}
types_clean <- all_types %>%
  filter(tbin == "t1") %>%
  mutate(gloss_clean = tolower(gloss))   %>%
  group_by(target_child_id, gloss_clean) %>%
  summarize(count = sum(count)) %>%
  filter(count >= MINWORDSFORVOCAB) %>%
  mutate(log_count = log(count)) %>%
  select(-count) %>%
  left_join(freq, by= c("gloss_clean" = "word")) %>%
  mutate(log_count_w1 = log_count,
         log_count_w2 = log_count,
         log_freq_w1 = log_freq, 
         log_freq_w2 = log_freq) %>%
  select(-log_count, -log_freq)
```

```{r}
conc <- read_csv("/Users/mollylewis/Documents/research/Projects/2_published/ref_complex/corpus/brysbaert_database/brysbaert_corpus.csv") %>%
  select(Word, Conc.M)

distinct_types <- types_clean %>%
  left_join(conc, by =c("gloss_clean" = "Word")) %>%
  ungroup() %>%
  distinct(gloss_clean, .keep_all = T)

median_conc <- median(distinct_types$Conc.M, na.rm = T)

types_clean_low <- types_clean %>%
    left_join(conc, by = c("gloss_clean" = "Word")) %>%
    filter(Conc.M < median_conc)

types_clean_high <- types_clean %>%
    left_join(conc, by = c("gloss_clean" = "Word")) %>%
    filter(Conc.M >= median_conc)
```

 
Get distance measure by kid at t1 in both wikipedia model and adult childes model
```{r}
get_vocab_measure_by_kid3 <- function(id, data, model){
  this_kids_model <- model %>%
    filter(target_word %in% data$gloss_clean)
  
  words_in_model <- data %>%
    filter(gloss_clean %in% this_kids_model$target_word)

   # get pairwise distances
   word_word_dists <- coop::cosine(t(this_kids_model[,-1])) 
   
     data.frame(target_child_id = id,
                mean_dist_t1 = mean(word_word_dists),
                median_dist_t1 = median(word_word_dists),
                var_dist_t1 =  ifelse(mean(var(word_word_dists)) == 0, NA,
                                       mean(var(word_word_dists))),
                n_t1 = nrow(word_word_dists),
                median_freq_t1 =  median(words_in_model$log_freq_w1, na.rm = T)) 
    
}

# low
nested_data_by_kid_low <- nest(types_clean_low, -target_child_id)
vocab_measures_childes_low <- map2_df(nested_data_by_kid_low$target_child_id, 
                          nested_data_by_kid_low$data, 
                          get_vocab_measure_by_kid3, 
                          word2vec_model_childes) %>%
  rename(mean_dist_t1_childes_low = mean_dist_t1,
         median_freq_t1_childes_low = median_freq_t1)

#high
nested_data_by_kid_high <- nest(types_clean_high, -target_child_id)
vocab_measures_childes_high <- map2_df(nested_data_by_kid_high$target_child_id, 
                          nested_data_by_kid_high$data, 
                          get_vocab_measure_by_kid3, 
                          word2vec_model_childes) %>%
  rename(mean_dist_t1_childes_high = mean_dist_t1,
         median_freq_t1_childes_high = median_freq_t1)

vocab_measures <- left_join(vocab_measures_childes_low, vocab_measures_childes_high, by = "target_child_id")  %>%
  mutate(mean_dist_t1_childes_log_low = log(mean_dist_t1_childes_low),
         mean_dist_t1_childes_log_high = log(mean_dist_t1_childes_high)) %>%
  select(-mean_dist_t1_childes_low, -mean_dist_t1_childes_high)
```

## Other kid-level variables at t1

Get merged dataframe
```{r}
vocab_df <- vocab_measures %>%
    left_join(kid_info %>% select(target_child_id, 
                                  log_mtld_t1, 
                                  log_mtld_t2, 
                                  age_t1, age_diff , 
                                  log_transcript_length_t1, log_transcript_length_t2,  
                                  log_num_trigrams_t1, log_num_trigrams_t2,
                                  mlu_m_t1, mlu_m_t2, 
                                  target_child_sex,
                                  conc_t1,
                                #  n_clusters, 
                                 # mean_trans_prob_t1 , 
                                 # prop_na_trans_t1,
                                  #Per_Corred_Pairs_No_Tax,
                                  Num_Corred_Pairs_No_Tax, 
                                  prop_noun_t1, 
                                  prop_verb_t1,
                                  corpus_id, collection_name,
                                age_t2)) 


```



## Predicting MTLD at t2 - childes
#### Mean dist and num corred
```{r}
low_mod <- lme4::lmer(scale(log_mtld_t2) ~ 
                  scale(mean_dist_t1_childes_log_low) + 
                  #scale(Num_Corred_Pairs_No_Tax) +
                  scale(age_t1) + 
                  scale(prop_noun_t1) +
                  scale(median_freq_t1_childes_low) +
                  scale(log_mtld_t1) +
                  scale(mlu_m_t1) +
                  scale(log_transcript_length_t1) + 
             (1|corpus_id),
   data = vocab_df) 

summary(low_mod)

high_mod <- lme4::lmer(scale(log_mtld_t2) ~ 
                  scale(mean_dist_t1_childes_log_high) + 
                  #scale(Num_Corred_Pairs_No_Tax) +
                  scale(age_t1) + 
                  scale(prop_noun_t1) +
                  scale(median_freq_t1_childes_high) +
                  scale(log_mtld_t1) +
                  scale(mlu_m_t1) +
                  scale(log_transcript_length_t1) + 
             (1|corpus_id),
   data = vocab_df) 

summary(high_mod)
```

```{r}
conf_low <- low_mod %>%
  confint %>%
  as.data.frame %>%
  rownames_to_column %>%
  filter(rowname %in% c("scale(mean_dist_t1_childes_log_low)", "scale(Num_Corred_Pairs_No_Tax)")) 


ts_low <- summary(low_mod)$coefficients %>%
    as.data.frame %>%
    rownames_to_column %>%
    filter(rowname %in% c("scale(mean_dist_t1_childes_log_low)", "scale(Num_Corred_Pairs_No_Tax)")) %>%
  left_join(conf_low) %>%
  mutate(group = "low")

conf_high<- high_mod %>%
  confint %>%
  as.data.frame %>%
  rownames_to_column %>%
  filter(rowname %in% c("scale(mean_dist_t1_childes_log_high)", "scale(Num_Corred_Pairs_No_Tax)")) 


ts_high<- summary(high_mod)$coefficients %>%
    as.data.frame %>%
    rownames_to_column %>%
    filter(rowname %in% c("scale(mean_dist_t1_childes_log_high)", "scale(Num_Corred_Pairs_No_Tax)")) %>%
  left_join(conf_high) %>%
  mutate(group = "high")

t_df <- bind_rows(ts_low, ts_high) %>%
  mutate(variable = c("log_mean_dist_t1",
                      "Num_Corred_Pairs_No_Tax",
                      "log_mean_dist_t1",
                      "Num_Corred_Pairs_No_Tax")) %>%
  rename(low_ci = `2.5 %`, 
         high_ci = `97.5 %`)

ggplot(t_df, aes(y = Estimate, fill = group, group = group, x = variable)) +
  xlab("Concreteness group") +
  geom_bar(stat = "identity",   position="dodge") +
  geom_linerange(aes(ymin = low_ci, 
                     ymax = high_ci),
                position=position_dodge(0.9))
  


