---
title: MTLD words by usefuleness
subtitle: defined as trigrams from cocoa
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: tango
    toc: true
    number_sections: false
    theme: cerulean
    toc_float: false
    code_folding: hide
---

```{r setup, message = F, warning = F}
library(knitr)

opts_chunk$set(echo = T, message = F, warning = F, 
               error = F, cache = F, tidy = F)

library(tidyverse)
library(langcog)
library(data.table)
library(feather)
library(broom)

theme_set(theme_classic(base_size = 10))
```

Do kids that know words at T2 (using mtld measure) know words that are more "useful" at T1? In particular, we define useful words to be words that allow you to make more trigrams in adult speech using coca spoken corpus (https://corpus.byu.edu/coca/).

FINDINGS:
  
  - no evidence for more probable trigrams for high vs. low kids at either group or individual level
  - Evidence that high kids have words for more frequent trigrams at group but not individual level
  - effect somewhat sensitive to number of words cutoff (effect bigger with fewer words)
  - but, lots of trigrams missing

## Group level {.tabset}

```{r}
# Get all unique words for two groups at t1. 
groups_info <- read_csv("../1_exploration/groups_info.csv")
target_types <- read_csv("../1_exploration/target_types_delta_450_1150.csv") %>%
  mutate(tbin = fct_recode(tbin, 
                           t1 = "low",
                           t2 = "high")) %>%
  mutate(gloss = tolower(gloss))

TOPNWORDS_PER_GROUP <- 500

t1_words <- target_types %>%
  filter(tbin == "t1") %>% # #t1 only
  left_join(groups_info %>% 
              mutate(target_child_id = as.numeric(target_child_id))) %>% # merge in group info
  group_by(delta_resid_group, gloss)  %>% # get glosses by group
  summarize(n = n()) %>% # get gloss counts by group
  top_n(TOPNWORDS_PER_GROUP) %>%
  group_by(delta_resid_group)  %>%
  slice(1:TOPNWORDS_PER_GROUP) %>%
  arrange(delta_resid_group, -n)
```

```{r}
coca_trigrams <- read_csv("coca_trigrams.csv") %>%
  select(w1, w2, w3, prob3, freq3) %>%
  data.table() 

t1_words_high <- t1_words %>%
                    filter(delta_resid_group == "high") %>%
                    select(gloss) %>%
  unlist(use.names = F)

t1_words_low <- t1_words %>%
                    filter(delta_resid_group == "low") %>%
                    select(gloss) %>%
  unlist(use.names = F)

high_trigrams <- coca_trigrams[w1 %in% t1_words_high &
                                 w2 %in% t1_words_high & 
                                 w3 %in% t1_words_high] %>%
  mutate(vocab_group = "high")

low_trigrams <- coca_trigrams[w1 %in% t1_words_low &
                                 w2 %in% t1_words_low & 
                                 w3 %in% t1_words_low] %>%
  mutate(vocab_group = "low")

all_trigrams <- bind_rows(high_trigrams, low_trigrams) %>%
  mutate(log_freq = log(freq3))
```

### Transitional probability
```{r}
ggplot(all_trigrams, aes(x = prob3, 
                         fill = vocab_group, 
                         group = vocab_group)) +
  geom_density(alpha = .4)

all_trigrams %>%
  group_by(vocab_group) %>%
  multi_boot_standard(col = "prob3", na.rm = T) %>%
  ggplot(aes(x = vocab_group, y = mean, group = vocab_group,
               fill = vocab_group)) +
  geom_bar(position = "dodge", stat = "identity") +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
                 position=position_dodge(width=0.9)) +
  theme_classic()
```

### Frequency
```{r}
ggplot(all_trigrams, aes(x = log_freq,  
                         fill = vocab_group, 
                         group = vocab_group)) +
  geom_density(alpha = .4)

all_trigrams %>%
  group_by(vocab_group) %>%
  multi_boot_standard(col = "log_freq", na.rm = T) %>%
    ggplot(aes(x = vocab_group, y = mean, group = vocab_group, 
               fill = vocab_group)) +
  geom_bar(position = "dodge", stat = "identity") +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
                 position=position_dodge(width=0.9)) +
  theme_classic()
```


## Individual Kids {.tabset}
### Transitional probability
```{r}
good_types_t1 <- target_types %>%
  left_join(groups_info %>% 
              mutate(target_child_id = as.numeric(target_child_id))) %>%
  filter(tbin == "t1")

get_trigrams_by_kid <- function(df, all_trigrams){
     current_trigrams <- all_trigrams[w1 %in% df$gloss &
                                 w2 %in% df$gloss & 
                                 w3 %in% df$gloss] 
     
     mean(current_trigrams$prob3, na.rm = T)
  
}

group_ids <-  good_types_t1 %>%
  distinct(target_child_id, 
         delta_resid_group)

trigram_freqs_by_kid <- good_types_t1 %>%
              nest(-target_child_id, -tbin) %>%
              mutate(mean_trigram_probability = 
                       map(data, get_trigrams_by_kid,
                           coca_trigrams)) %>%
            select(-data) %>%
            unnest() %>%
            left_join(groups_info %>% select(target_child_id, delta_resid_group)) 


ggplot(trigram_freqs_by_kid, 
       aes(x = mean_trigram_probability, 
                         fill = delta_resid_group, 
                         group = delta_resid_group)) +
  geom_density(alpha = .4)


trigram_freqs_by_kid %>%
  group_by(delta_resid_group) %>%
  multi_boot_standard(col = "mean_trigram_probability",
                      na.rm = T) %>%
    ggplot(aes(x = delta_resid_group, y = mean, group = delta_resid_group, fill = delta_resid_group)) +
  geom_bar(position = "dodge", stat = "identity") +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
                 position=position_dodge(width=0.9)) +
  theme_classic()

```


### Frequency
```{r}
good_types_t1 <- target_types %>%
  left_join(groups_info %>% 
              mutate(target_child_id = as.numeric(target_child_id))) %>%
  filter(tbin == "t1")

get_trigrams_by_kid <- function(df, all_trigrams){
     current_trigrams <- all_trigrams[w1 %in% df$gloss &
                                 w2 %in% df$gloss & 
                                 w3 %in% df$gloss] 
     
     mean(log(current_trigrams$freq3), na.rm = T)
  
}

group_ids <-  good_types_t1 %>%
  distinct(target_child_id, 
         delta_resid_group)

trigram_freqs_by_kid <- good_types_t1 %>%
              nest(-target_child_id, -tbin) %>%
              mutate(log_mean_trigram_freq= 
                       map(data, get_trigrams_by_kid,
                           coca_trigrams)) %>%
            select(-data) %>%
            unnest() %>%
            left_join(groups_info %>% select(target_child_id, delta_resid_group))


ggplot(trigram_freqs_by_kid, 
       aes(x = log_mean_trigram_freq, 
                         fill = delta_resid_group, 
                         group = delta_resid_group)) +
  geom_density(alpha = .4)


trigram_freqs_by_kid %>%
  group_by(delta_resid_group) %>%
  multi_boot_standard(col = "log_mean_trigram_freq",
                      na.rm = T) %>%
  ggplot(aes( x = delta_resid_group, y = mean, group = delta_resid_group, fill = delta_resid_group)) +
  geom_bar(position = "dodge", stat = "identity") +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
                 position=position_dodge(width=0.9)) +
  theme_classic()

```
