---
title: MTLD words by usefuleness
subtitle: defined as adult trigrams from childes
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: tango
    toc: true
    number_sections: false
    theme: cerulean
    toc_float: false
    code_folding: hide
---

```{r setup, message = F, warning = F}
library(knitr)

opts_chunk$set(echo = T, message = F, warning = F, 
               error = F, cache = F, tidy = F)

library(tidyverse)
library(langcog)
library(data.table)
library(feather)
library(broom)

theme_set(theme_classic(base_size = 10))
```

Do kids that know more words at T2 (using mtld measure) know words that are more "useful" at T1? In particular, we define useful words to be words that allow you to make more trigrams in adult speech from childes (child-directed). Because high kids have bigger vocabs, we sample TOPNWORDS_PER_GROUP.

FINDINGS:



## Group level  {.tabset}

```{r}
MINCOUNT <- 0
TOPNWORDS_PER_GROUP <- 500

# Get all unique words for two groups at t1. 
groups_info <- read_csv("../1_mtld_measure/data/groups_info_600_900.csv")
target_types <- read_csv("../1_mtld_measure/data/target_types_for_MTLD_kids_600_900.csv") %>%
  group_by(target_child_id, tbin, gloss) %>%
  summarize(count = sum(count)) %>%
  ungroup() %>%
  filter(count >= MINCOUNT) %>%
  mutate(gloss = tolower(gloss))


t1_words <- target_types %>%
  filter(tbin == "t1") %>% # #t1 only
  left_join(groups_info %>% 
              mutate(target_child_id = as.numeric(target_child_id))) %>% # merge in group info
  group_by(delta_resid_group, gloss)  %>% # get glosses by group
  summarize(n = n()) %>% # get gloss counts by group
  top_n(TOPNWORDS_PER_GROUP) %>%
  group_by(delta_resid_group)  %>%
  slice(1:TOPNWORDS_PER_GROUP) %>%
  arrange(delta_resid_group, -n)
```


```{r}
childes_trigrams <- read_csv("data/trigrams/adult_childes_trigrams_turns.csv") %>%
  data.table() 

t1_words_high <- t1_words %>%
  ungroup()%>%
                    filter(delta_resid_group == "high") %>%
                    select(gloss) %>%
  unlist(use.names = F)


t1_words_low <- t1_words %>%
    ungroup() %>%
                    filter(delta_resid_group == "low") %>%
                    select(gloss) %>%
  unlist(use.names = F)

high_trigrams <- childes_trigrams[w1 %in% t1_words_high &
                                 w2 %in% t1_words_high & 
                                 w3 %in% t1_words_high] %>%
  mutate(vocab_group = "high")

low_trigrams <- childes_trigrams[w1 %in% t1_words_low &
                                 w2 %in% t1_words_low & 
                                 w3 %in% t1_words_low] %>%
  mutate(vocab_group = "low")

all_trigrams <- bind_rows(high_trigrams, low_trigrams)
```

### number of trigrams
```{r}
count(all_trigrams, vocab_group) %>%
  kable()
```

### frequency of trigrams
```{r}
all_trigrams %>%
  mutate(log_freq = log(freq)) %>%
  group_by(vocab_group) %>%
  multi_boot_standard(col = "log_freq", na.rm = T) %>%
  ggplot(aes( x = vocab_group, y = mean, group = vocab_group, fill = vocab_group)) +
  geom_bar(position = "dodge", stat = "identity") +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
                 position=position_dodge(width=0.9)) +
  theme_classic()

  #summarize(mean_freq= mean(log_freq),
  #          ci_lower_freq = mean(log_freq) - (1.96 * sd(log_freq)),
  #          ci_upper_freq = mean(log_freq) + (1.96 * sd(log_freq)))
  

 all_trigrams %>%
  group_by(vocab_group) %>%
  multi_boot_standard(col = "freq", na.rm = T) %>%
  kable()

```


## Individual Kids  {.tabset}
```{r}
good_types_t1 <- target_types %>%
  left_join(groups_info %>% 
              mutate(target_child_id = as.numeric(target_child_id))) %>%
  filter(tbin == "t1") %>%
  filter(gloss %in% t1_words$gloss)

get_trigrams_by_kid <- function(df, all_trigrams, measure){
     current_trigrams <- all_trigrams[w1 %in% df$gloss &
                                 w2 %in% df$gloss & 
                                 w3 %in% df$gloss] 
     
     if (measure == "num"){
      nrow(current_trigrams)
     } else if (measure == "freq"){
      mean(log(current_trigrams$freq))
     }
}
```

### Number

```{r}
trigram_num_by_kid <- good_types_t1 %>%
              nest(-target_child_id, -tbin) %>%
              mutate(trigram_num = 
                       map(data, get_trigrams_by_kid,
                           childes_trigrams, "num")) %>%
            select(-data) %>%
            unnest() %>%
            left_join(groups_info %>% select(target_child_id, delta_resid_group)) %>%
            mutate(log_trigram_num = log(trigram_num)) 



ggplot(trigram_num_by_kid, 
       aes(x = log_trigram_num, 
                         fill = delta_resid_group, 
                         group = delta_resid_group)) +
  geom_density(alpha = .4) +
  theme_classic()


ggplot(trigram_num_by_kid, 
       aes(x = log_trigram_num, 
                         fill = delta_resid_group)) +
       geom_histogram() +
  theme_classic()

trigram_num_by_kid %>%
  group_by(delta_resid_group) %>%
  multi_boot_standard(col = "log_trigram_num", na.rm = T) %>%
    ggplot(aes( x = delta_resid_group, y = mean, 
                group = delta_resid_group, fill = delta_resid_group)) +
  geom_bar(position = "dodge", stat = "identity") +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
                 position=position_dodge(width=0.9)) +
  theme_classic()
```


### Frequency
```{r}
trigram_freqs_by_kid <- good_types_t1 %>%
              nest(-target_child_id, -tbin) %>%
              mutate(log_mean_trigram_frequency = 
                       map(data, get_trigrams_by_kid,
                           childes_trigrams, "freq")) %>%
            select(-data) %>%
            unnest() %>%
            left_join(groups_info %>% select(target_child_id, delta_resid_group))


ggplot(trigram_freqs_by_kid, 
       aes(x = log_mean_trigram_frequency, 
                         fill = delta_resid_group, 
                         group = delta_resid_group)) +
  geom_density(alpha = .4) +
  theme_classic()


trigram_freqs_by_kid %>%
  group_by(delta_resid_group) %>%
  multi_boot_standard(col = "log_mean_trigram_frequency", na.rm = T) %>%
    ggplot(aes( x = delta_resid_group, y = mean, group = delta_resid_group, fill = delta_resid_group)) +
  geom_bar(position = "dodge", stat = "identity") +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), 
                 position=position_dodge(width=0.9)) +
  theme_classic()
```
