---
title: Semantic centrality analyses
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: tango
    toc: true
    number_sections: false
    theme: cerulean
    toc_float: false
    code_folding: hide
runtime: shiny
---

```{r setup, message = F, warning = F}
library(knitr)
library(tidyverse)
library(corrplot)
library(shiny)

opts_chunk$set(echo = T, message = F, warning = F, 
               error = F, cache = F, tidy = F)


theme_set(theme_classic(base_size = 15))
```

This is similiar to the previous analyses with a few modifications. 

* dropped density measure for simplification (it's highly correlated with centrality, and more complex)
* added measure of transcript length - turns out it's highly correlated with number of types measure, and linearly related
* added n kid trigrams measure

```{r}
vocab_df_trim <- read_csv("semantic_density_tidy_df_MIN1.csv")
```

## Pairwise correlations
```{r, fig.width = 10, fig.height = 10}
ALPHA <- .05
corr_mat <- cor(vocab_df_trim, 
                use = "pairwise.complete.obs")

p.mat <- cor.mtest(vocab_df_trim, 
                  conf.level = (1-ALPHA),  
                  use = "pairwise.complete.obs")$p

cols = rev(colorRampPalette(c("red", "white", "blue"))(100))

corrplot(corr_mat, method = "color",  col = cols,
         type = "upper", order = "original", number.cex = .7,
         addCoef.col = "black", 
         p.mat = p.mat, sig.level = ALPHA, insig = "blank", 
         tl.col = "black", tl.srt = 90,
         diag = FALSE)
```


### Explore correlations between measures
```{r}
shinyApp(

  ui = fluidPage(
    tags$head(tags$style(type = "text/css", '.well{width: 520px}')),
        selectInput("xvar", "X variable",  names(vocab_df_trim), selected = "log_n_parent_trigrams_t1"),
        selectInput("yvar", "Y variable",  names(vocab_df_trim), selected = "log_n_parent_trigrams_t2"),
    plotOutput("scatter_plot")
  ),
  
  server = function(input, output) {
    
    output$scatter_plot <- renderPlot({
      ggplot(vocab_df_trim, aes_string(input$xvar, input$yvar)) +
        geom_point() +
        geom_smooth(alpha = .4) +
        geom_smooth(method = "lm", color = "red") 

    })
  },
  
  options = list(height = 600)
)

```


## Regressions
### Predicting log_n_kid_trigrams_t2
```{r}
lm(log_n_kid_trigrams_t2 ~ 
     log_centrality_t1 +
     age_diff +
     log_n_kid_trigrams_t1 +
     mean_log_word_freq_t1 + 
     mean_log_word_freq_t2 +
     log_transcript_length_t1 +
     log_transcript_length_t2,
     vocab_df_trim) %>%
  summary()

lm(log_n_kid_trigrams_t2 ~ 
     log_centrality_t1 +
     age_diff +
     log_n_kid_trigrams_t1 +
     mean_log_word_freq_t1 + 
     mean_log_word_freq_t2 +
     log_transcript_length_t1,
     #log_transcript_length_t2,
     vocab_df_trim %>% filter(log_centrality_t1 > -1.95)) %>%
  summary()
```

### Predicting log_n_parent_trigrams_t2
```{r}
lm(log_n_parent_trigrams_t2 ~ 
     log_centrality_t1 +
     age_diff +
     log_n_parent_trigrams_t1 +
     log_transcript_length_t1 +
     mean_log_word_freq_t1   +
     mean_log_word_freq_t2,
     vocab_df_trim) %>%
  summary()

lm(log_n_parent_trigrams_t2 ~ 
     log_centrality_t1 +
     age_diff +
     log_n_parent_trigrams_t1 +
     log_transcript_length_t1 +
     mean_log_word_freq_t1   +
     mean_log_word_freq_t2,
     vocab_df_trim %>% filter(log_centrality_t1 > -1.95)) %>%
  summary()

lm(log_n_parent_trigrams_t2 ~ 
     log_centrality_t1 +
     age_diff +
     log_n_parent_trigrams_t1 +
     log_transcript_length_t1 +
     mean_log_word_freq_t1   +
     mean_log_word_freq_t2,
     vocab_df_trim %>% filter(log_centrality_t1 > -1.95)) %>%
  summary()


```

### Predicting log_mtld_t2
```{r}
lm(log_mtld_t2 ~ 
     log_centrality_t1+
     age_diff +
     log_transcript_length_t1 +
     log_mtld_t1,
     vocab_df_trim) %>%
  summary()

lm(log_mtld_t2 ~ 
     log_centrality_t1+
     age_diff +
     log_mtld_t1,
     vocab_df_trim %>% filter(log_centrality_t1 > -1.95)) %>%
  summary()


vocab_df_trim %>%  
  lm(formula = log_n_kid_trigrams_t2 ~ 
       log_centrality_t1 +
       age_diff +
       log_n_kid_trigrams_t1 +
       mean_log_word_freq_t1 +
       mean_log_word_freq_t2, data = .) %>% 
  #summary() %>%
  car::vif()

```