---
title: MTLD words by usefuleness
subtitle: defined as adult trigrams
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: tango
    toc: true
    number_sections: false
    theme: cerulean
    toc_float: false
    code_folding: hide
---

```{r setup, message = F, warning = F}
library(knitr)

opts_chunk$set(echo = T, message = F, warning = F, 
               error = F, cache = F, tidy = F)

library(tidyverse)
library(langcog)
library(data.table)
library(feather)
library(broom)

theme_set(theme_classic(base_size = 10))
```

Do kids that know words at T2 (using mtld measure) know words that are more "useful" at T1? In particular, we define useful words to be words that allow you to make more trigrams in adult speech using coca spoken corpus (https://corpus.byu.edu/coca/).

FINDINGS:
  
  - no evidence for more probable trigrams for high vs. low kids at either group or individual level
  - but, lots of trigrams missing

## Group level

```{r}
# Get all unique words for two groups at t1. 
groups_info <- read_csv("groups_info.csv")
target_types <- read_csv("target_types_delta_450_1150.csv") %>%
  mutate(tbin = fct_recode(tbin, 
                           t1 = "low",
                           t2 = "high")) %>%
  mutate(gloss = tolower(gloss))

MIN_KIDS_PRODUCED_WORD <- 5

good_words <- target_types %>%
  count(gloss, target_child_id) %>% # one row a word-kid
  count(gloss) %>%
  filter(nn >= MIN_KIDS_PRODUCED_WORD) # more than n words?

good_kids <- target_types %>%
  filter(gloss %in% good_words$gloss) %>%
  count(target_child_id, tbin) %>% # each row a tbin-kid
  count(target_child_id) %>%
  filter(nn > 1) # two time points?

# get the good words, filtering on kids and words
good_types <- target_types %>%
    filter(gloss %in% good_words$gloss,
           target_child_id %in% good_kids$target_child_id)

# get t1 words only
t1_words <- good_types %>%
  left_join(groups_info %>% 
              mutate(target_child_id = as.numeric(target_child_id))) %>%
  filter(tbin == "t1") %>%
  distinct(tbin, delta_resid_group, gloss)
```

```{r}
coca_trigrams <- read_csv("results_freq_prob_coca.csv") %>%
  select(w1, w2, w3, prob3, prob2_1) %>%
  data.table() 

t1_words_high <- t1_words %>%
                    filter(delta_resid_group == "high") %>%
                    select(gloss) %>%
  unlist(use.names = F)

t1_words_low <- t1_words %>%
                    filter(delta_resid_group == "low") %>%
                    select(gloss) %>%
  unlist(use.names = F)

high_trigrams <- coca_trigrams[w1 %in% t1_words_high &
                                 w2 %in% t1_words_high & 
                                 w3 %in% t1_words_high] %>%
  mutate(vocab_group = "high")

low_trigrams <- coca_trigrams[w1 %in% t1_words_low &
                                 w2 %in% t1_words_low & 
                                 w3 %in% t1_words_low] %>%
  mutate(vocab_group = "low")

all_trigrams <- bind_rows(high_trigrams, low_trigrams)

ggplot(all_trigrams, aes(x = prob3, 
                         fill = vocab_group, 
                         group = vocab_group)) +
  geom_histogram()

all_trigrams %>%
  group_by(vocab_group) %>%
  multi_boot_standard(col = "prob3", na.rm = T)

```


## Individual Kids
```{r}
good_types_t1 <- good_types %>%
  left_join(groups_info %>% 
              mutate(target_child_id = as.numeric(target_child_id))) %>%
  filter(tbin == "t1")

get_trigrams_by_kid <- function(df, all_trigrams){
     current_trigrams <- all_trigrams[w1 %in% df$gloss &
                                 w2 %in% df$gloss & 
                                 w3 %in% df$gloss] 
     
     mean(current_trigrams$prob3, na.rm = T)
  
}

group_ids <-  good_types_t1 %>%
  distinct(target_child_id, 
         delta_resid_group)

trigram_freqs_by_kid <- good_types_t1 %>%
              nest(-target_child_id, -tbin) %>%
              mutate(mean_trigram_frequency = 
                       map(data, get_trigrams_by_kid,
                           coca_trigrams)) %>%
            select(-data) %>%
            unnest() %>%
            left_join(group_ids)


ggplot(trigram_freqs_by_kid, 
       aes(x = mean_trigram_frequency, 
                         fill = delta_resid_group, 
                         group = delta_resid_group)) +
  geom_density(alpha = .4)


trigram_freqs_by_kid %>%
  group_by(delta_resid_group) %>%
  multi_boot_standard(col = "mean_trigram_frequency",
                      na.rm = T)

```
